---
title: "Taking advantage of all the symbols in ggplot2"
layout: page
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
permalink: /symbols.html
header:
  overlay_color: '444444'
editor_options: 
  chunk_output_type: console
---

There are a few things in R that are particularly annoying and I have left them all to the last steps. This is the case for any package or set of tools we choice to use as researchers. The issues below are a random collection of issues do with the tidyverse approach and the other packages I use to do analysis. Here are my learning steps for the annoying R issues.

# `ggplot2` symbols

```{r global-options, message=FALSE, warning=FALSE}
# export .r code only
# knitr::purl("./Davidson_2019_BeechForest.Rmd")

# render draft to webpage
# rmarkdown::render(input = "Davidson_2019_BeechForest.Rmd")
# ,
#                   output_format = "html_document",
#                   output_file = "Davidson_2019_t.html")

#document global rules
knitr::opts_chunk$set(comment=NA,
                      fig.path = "./figs/",
                      echo=FALSE, 
                      include = FALSE,
                      fig.height=5.5, 
                      fig.width=12.5,
                      message=FALSE, 
                      warning=FALSE, echo=FALSE)
# how do I do this??
# ,eval = FALSE,include = FALSE

# libraries needed
source("./rcode/r-packages-needed.R", echo = FALSE)

# themes
# source("./Rcode/davidson-2019-theme.r", echo = FALSE)

#overall code
# source("./rcode/manuscript-source-code.R", echo = FALSE)
```


```{r plot-data}

# import study design data
# find code in script plot-study-data.R
# bbech-forest-private repo
# can use "source()" to add this

plot.dat.all1 <- read_csv("./data/study-design-plot-input-data.csv") %>%
                    mutate(Date = true.date)

```

I have been writing my first paper for my PhD for a ridiculously long time and as I am finally finishing figures for the manuscript I have had huge issues trying to sort these symbols.

## My solution for this script

### `aes(shape = Valley; colour = Control, fill = Rats)`

Overll the three factors ploted as colours looks like so:

```{r plot-1, echo=FALSE, message=FALSE, warning=FALSE, include = TRUE}
p1 <- ggplot(plot.dat.all1, aes(y = Treatments, 
                          x = Date)) +
  geom_line() +
  geom_point(aes(col = Control),
             size = 3, 
             stroke = 1.2)

p2 <- ggplot(plot.dat.all1, aes(y = Treatments, 
                          x = Date)) +
  geom_line() +
  geom_point(aes(col = Rats),
             size = 3, 
             stroke = 1.2)

p3 <- ggplot(plot.dat.all1, aes(y = Treatments, 
                                x = Date)) +
  geom_line() +
  geom_point(aes(col = Valley),
             size = 3, 
             stroke = 1.2)



#easy plotting
gridExtra::grid.arrange(p1,p2,p3)

# grobs save?
# plots3 <- grid.arrange(p1,p2,p3)

# export plot for example vignette
# jpeg("./figs/plot3-symbols.jpeg")
# grid.arrange(p1,p2,p3)
# dev.off()
```

[Figure 1: Plot of the three different factors, each with two levels that I want to each have the following representative symbols](./figs/plot3-symbols.jpeg)

```{r current sampling allociation}

```

For the example above there are three factors (valley, control, rats) each has 2 levels. However, because we are missing some combinations we have a total of 6 treatments.

```{r finding-all-combos}
vars.names <- c("valley", "control", "rats")
```

## `aes(shape = Valley; colour = Control)`

To plot this I need to split the symbols. I have decided to do this as follows: `shape = Valley`; `colour = Control`; `fill = Rats`. And in ggplot2 this is wrapped with the `aes()` function.

```{r plot2-valley-shape, include = TRUE}
p1.1 <- ggplot(plot.dat.all1, aes(y = Treatments, 
                          x = Date)) +
  geom_line() +
  geom_point(aes(col = Control, 
                 shape = Valley),
             size = 3, 
             stroke = 1.2)

p2.1 <- ggplot(plot.dat.all1, aes(y = Treatments, 
                          x = Date)) +
  geom_line() +
  geom_point(aes(col = Rats, 
                 shape = Valley),
             size = 3, 
             stroke = 1.2)

#easy plotting
gridExtra::grid.arrange(p1.1,p2.1)

# export plot for example vignette
# jpeg("./figs/plot2-symbols.jpeg")
# grid.arrange(p1.1,p2.1)
# dev.off()
```

[Figure 2: PLots including shape for Valley but two different plots showing the different factors in Control and Rats factors](./figs/plot2-symbols.jpeg)

## `aes(shape = Valley; colour = Control, fill = Rats)`

And in ggplot2 this is wrapped with the `aes(shape = Valley; colour = Control, fill = Rats)` function. And then it looks like this:

```{r plot1-valley-contol-rats, include = TRUE}
p1.2 <- ggplot(plot.dat.all1, aes(y = Treatments, 
                          x = Date)) +
  geom_line() +
  geom_point(aes(col = Control, 
                 shape = Valley, fill = Rats),
             size = 3,alpha = 0.2 ,
             stroke = 1.2) +
  theme(legend.position = "none")

p2.2 <- ggplot(plot.dat.all1, aes(y = Treatments, 
                          x = Date)) +
  geom_line() +
  geom_point(aes(col = Rats, 
                 shape = Valley, fill = Rats),
             size = 3, 
             stroke = 1.2) +
  theme(legend.position = "none")

p1.2

#easy plotting
# gridExtra::grid.arrange(p1.2,p2.2)

# export plot for example vignette
# jpeg("./figs/plot1alpha-symbols.jpeg")
# p1.2
# dev.off()
```

[Figure 3: Attempting to understand why the colouring is not working with fill](./figs/plot1alpha-symbols.jpeg)
 
**WHAT makes this such a hard question??**

But for some reason fill doesn't work the way I want with rings for rats! 

## Working out plots space

Working with factors below to check this is not the issue.

```{r data-symbol-issues}

#making data prefect again
plot.dat.all2 <-
  plot.dat.all1 %>%
  select(Valley, Control, Rats, Date, Treatments) %>%
    mutate(yyDate = strftime(Date, "%Y"))
   
plot.dat.all2$Valley
str(plot.dat.all2$Valley)      
levels(plot.dat.all2$Valley)
labels(plot.dat.all2$Valley)

plot.dat.all2$Control
str(plot.dat.all2$Control) 
levels(plot.dat.all2$Control)
labels(plot.dat.all2$Control)

plot.dat.all2$Rats
str(plot.dat.all2$Rats)  
levels(plot.dat.all2$Rats)
labels(plot.dat.all2$Rats)

table(plot.dat.all2$Valley, plot.dat.all2$Control, plot.dat.all2$Rats)
```

IN this case the issue is exactly that to do with the 8 grids and the 6 treatments across 20 time intervals. 

```{r not-working-plot}

ggplot(plot.dat.all2, aes(y = Treatments, 
                          x = yyDate)) +
  geom_line() +
  geom_jitter(aes(col = Control, 
                 shape = Valley, fill = Rats),
             size = 3,alpha = 0.2 ,
             stroke = 1.2) +
  theme(legend.position = "none")

```

- I am confused about these replicates but none are as complex is what I have here.

## Solutions

1. First issue is that it is by grid not treatment

```{r not-working-plot2}

ggplot(plot.dat.all1, aes(y = grid, 
                          x = Date)) +
  geom_line() +
  geom_point(aes(col = Control, 
                 shape = Valley), fill = "black",
             size = 3,alpha = 0.6,
             stroke = 5) +
  theme(legend.position = "none")

```

## This is why?? I had a look at examples I could find that have done this and the following stack-exchange and other blogs helped to produce this:

EXEPT ONE [here](http://www.audhalbritter.com/complicated-figure/)

or a few more options... 

2. And it turns out you can do this by creating a dumby variable

3. Or change alpha to 0.5 so solid lines come up? 

4. Or us alpha as facet

5. Change size of some just slightly so can see black?

```{r data-symbol-issues2}

#making data prefect again
plot.dat.all2 <-
  plot.dat.all1 %>%
  select(Valley, Control, Rats, Date, Treatments) %>%
    mutate(yyDate = strftime(Date, "%Y"))
   
plot.dat.all2$Valley
str(plot.dat.all2$Valley)      
levels(plot.dat.all2$Valley)
labels(plot.dat.all2$Valley)

plot.dat.all2$Control
str(plot.dat.all2$Control) 
levels(plot.dat.all2$Control)
labels(plot.dat.all2$Control)

plot.dat.all2$Rats
str(plot.dat.all2$Rats)  
levels(plot.dat.all2$Rats)
labels(plot.dat.all2$Rats)

table(plot.dat.all2$Valley, plot.dat.all2$Control, plot.dat.all2$Rats)
```


```{r}
p.design <- plot.dat.all1 %>%
              mutate(grid = as.numeric(factor(grid)))

glimpse(plot.dat.all1)
```



```{r working-plot1,include = TRUE}

source("./rcode/davidson_2019_theme.r")

ggplot(p.design, aes(y = factor(grid),col = Rats, 
                          shape = Valley, 
                          fill = Control, 
                          x = Date,
                     group = grid), size = 4) +
 geom_line(col = "grey50") +
 geom_point(aes(size = Rats), stroke = 1.25) +
  
  scale_color_manual(name = "Stoat Control", 
                     values = c("white", "black")) +
  
  scale_shape_manual(name = "Ecosystem", 
                     values = c(24, 21)) +
  
  # manually define the fill colours
  
  scale_fill_manual(name = "Stoat Control", 
                     values = c("lightblue", "darkorange")) +
# theme
theme_new() +

  # labels
# scale_x_date() +
# defining size with 2 marginally different values
  scale_size_manual(name = "Rat Treatment", values = c(4, 3)) +

# Remove fill legend and replace the fill legend using the newly created size
  guides(col = "none", 
         size = guide_legend(override.aes = list(shape = c(16, 1))),
         shape = guide_legend(override.aes = list(shape = c(24, 21), size = 4)),
         fill = guide_legend(override.aes = list(col = c("lightblue", "darkorange"), size = 4))) +
  xlab("Timing of sample") +
  ylab("Grid Location")

```

## Resources

Dealing with factors can be hard in R and here are some resources for FACTORS in R:

- [A collection of scripts I use](https://github.com/davan690/usefulr/)

- [An .Rmd file for factoring](https://github.com/davan690/usefulr/tree/gh-pages/RMarkdown-vigettes)

- Simple example 1

```{r option1}

x<-c(2.2, 3, 3.8, 4.5, 7, 8.5, 6.7, 5.5)
y<-c(4, 5.5, 4.5, 9, 11, 15.2, 13.3, 10.5)

# Plot points
plot(x, y)
# Change plotting symbol
# Use solid circle
plot(x, y, pch = 19)

# what about making sure shape is fill--possible
# Fill shapes are:

# Change color
plot(x, y, pch=19, col="darkgreen", cex=1.5)

# Color can be a vector
plot(x, y, pch=19, col=c("green", "red"))

# change border, background color and line width
plot(x, y, pch = 24, cex=2, col="blue", bg="red", lwd=2)



# Another option
df <- mtcars[, c("mpg", "cyl", "wt")]
df$cyl <- as.factor(df$cyl)
head(df)
# Scatter plot with multiple groups
# shape depends on cyl
ggplot(df, aes(x=wt, y=mpg, group=cyl)) +
  geom_point(aes(shape=cyl))
# Change point shapes and colors
ggplot(df, aes(x=wt, y=mpg, group=cyl)) +
  geom_point(aes(shape=cyl, color=cyl))
# change point shapes,  colors and sizes
ggplot(df, aes(x=wt, y=mpg, group=cyl)) +
  geom_point(aes(shape=cyl, color=cyl, size=cyl))
# Change colors and shapes manually
ggplot(df, aes(x=wt, y=mpg, group=cyl)) +
  geom_point(aes(shape=cyl, color=cyl), size=2)+
  scale_shape_manual(values=c(3, 16, 17))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9'))+
  theme(legend.position="top")
# Change the point size manually
ggplot(df, aes(x=wt, y=mpg, group=cyl)) +
  geom_point(aes(shape=cyl, color=cyl, size=cyl))+
  scale_shape_manual(values=c(3, 16, 17))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9'))+
  scale_size_manual(values=c(2,3,4))+
  theme(legend.position="top")
```

- Simple example 2

```{r option2}

x<-c(2.2, 3, 3.8, 4.5, 7, 8.5, 6.7, 5.5)
y<-c(4, 5.5, 4.5, 9, 11, 15.2, 13.3, 10.5)
# Plot points
plot(x, y)
# Change plotting symbol
# Use solid circle
plot(x, y, pch = 19)

# what about making sure shape is fill--possible
# Fill shapes are:

# Change color
plot(x, y, pch=19, col="darkgreen", cex=1.5)

# Color can be a vector
plot(x, y, pch=19, col=c("green", "red"))

# change border, background color and line width
plot(x, y, pch = 24, cex=2, col="blue", bg="red", lwd=2)

```

- Complex fill solution
```{r complex-fix}
# create a data set
Data <- data_frame(Temperature = c(8.77, 8.67, 7.47, 7.58, 9.1, 8.9, 7.5, 7.7),
                   Precipitation = c(1848, 3029, 1925, 2725, 1900, 3100, 
                                     2000, 2800),
                   Temperature_level = as.factor(c(rep("subalpine", 2), 
                                                   rep("alpine", 2), 
                                                   rep("subalpine", 2), 
                                                   rep("alpine", 2))),
                   Precipitation_level = as.factor(c(rep(c(1,2),4))),
                   Source = c(rep("long term", 4), rep("short term", 4)))

p <- ggplot(Data, aes(x = Precipitation, 
                      y = Temperature, 
                      color = Precipitation_level, 
                      shape = Temperature_level, 
                      fill = factor(ifelse(Source == "short term", 
                                           Precipitation_level, Source)))) +
  
  scale_color_manual(name = "Precipitation level", 
                     values = c("skyblue1", "steelblue3")) +
  
  scale_shape_manual(name = "Temperature level", values = c(24, 21)) +
  
  # manually define the fill colours
  scale_fill_manual(name = "Source", 
                    values = c("skyblue1", "steelblue3", "white")) +
  theme_minimal()

p + geom_point(size = 3)


p + 
  # add size for Source
  geom_point(aes(size = Source)) +
  # defining size with 2 marginally different values
  scale_size_manual(name = "Source", values = c(3, 3.01)) +
  # Remove fill legend and replace the fill legend using the newly created size
  guides(fill = "none", 
         size = guide_legend(override.aes = list(shape = c(1, 16))))

```

# The dreaded legend

Legends are really tricky to work out in ggplot I find. Manly for one reason, that they are directly linked to the data rather than manually entering it. Anyway here is my code to manually control the legend.

```{r}
# By grid plot
# Base plot code
# dat.msr.1 comes from....

# # Can I turn this into a function I can use to select either one of 3 different estimates of abundance into
# # seed, density, rats?
# meanS <- dat.msr.1 %>%
#   filter(spp == "seed") %>%
#   mutate(true.date = as.Date(true.date),
#          mean.s = est.dat)
#
# #comes from summary Rscript.
# # source(./Rcode/figures/summar...)
# meanS.1 <- mean1 %>%
#   filter(spp == "seed")  %>%
#   mutate(true.date = as.Date(true.date))

# Cleaning up total dataset
plot.dat.all1 # already done!

# summaries
seed.mean <- plot.dat.all1 %>%
  group_by(Control, Valley, Date) %>%
  summarise(
    mean.s = mean(cum.seed),
    sd.s = sd(cum.seed),
    se.s = sd(cum.seed) / sqrt(length(cum.seed)) * 1.96,
    lcl.s = mean(cum.seed) - (sd(cum.seed) / sqrt(length(cum.seed)) *
                                1.96),
    ucl.s = mean(cum.seed) + (sd(cum.seed) / sqrt(length(cum.seed)) *
                                1.96)
  ) %>%
  ungroup()

seed.mean <- seed.mean %>%
  mutate(
    gp.treat = factor(paste(Valley, Control)),
    cum.seed = mean.s,
    Rats = factor("Full")
  )


levels(seed.mean$gp.treat)
# note that plot uses both meanS and meanS.1
# seed.aver.plot <-
ggplot(data = plot.dat.all1,
       aes(
         y = cum.seed,
         x = Date,
         shape = Valley,
         fill = Control,
         col = Rats
       )) +
  
  # extra goodies
  geom_rect(aes(
    xmin = ymd('2000-01-01'),
    xmax = ymd('2000-12-31'),
    ymin = -Inf,
    ymax = Inf
  ),
  colour = "grey90",
  fill = "grey90") +
  
  geom_rect(aes(
    xmin = ymd('2002-01-01'),
    xmax = ymd('2002-12-31'),
    ymin = -Inf,
    ymax = Inf
  ),
  colour = "grey90",
  fill = "grey90") +
  
  geom_rect(aes(
    xmin = ymd('2004-01-01'),
    xmax = ymd('2004-12-31'),
    ymin = -Inf,
    ymax = Inf
  ),
  colour = "grey90",
  fill = "grey90") +
  
  geom_point(aes(fill = Control), stroke = 1.15, size = 3) +
  
  #   geom_line(data = seed.mean, aes(y = mean.s,
  #                                   x = Date, group = gp.treat), size = 0.9) +
  #
  # # +
  #   # geom_errorbar(data = meanS.1, mapping = aes(ymin = lcl.s, ymax = ucl.s), width = 0, alpha = 0.5, lwd = 0.75) +
  #
  geom_line(data = seed.mean,
            aes(y = mean.s,
                x = Date),
            size = 1,
            col = "grey50") +
  
  geom_point(data = seed.mean, aes(y = mean.s,
                                   x = Date), size = 5) +
  
  
  
  scale_color_manual(name = "Stoat Control",
                     values = c("white", "black")) +
  
  scale_shape_manual(name = "Ecosystem",
                     values = c(24, 21)) +
  
  # manually define the fill colours
  #
  scale_fill_manual(
    name = "Stoat Control",
    values = c("cornflowerblue", "darkorange", "white", "black")
  ) +
  
  # defining size with 2 marginally different values
  scale_size_manual(name = "Rat Control", values = c(5, 4)) +
  
  # geom_abline(intercept = 5, slope = 0, size = 1) +
  # theme
  theme_new() +
  
  
  # labels
  # scale_y_discrete(labels = c("1","2","3","4"," ","1","2","3","4")) +
  # scale_x_date() +
  # defining size with 2 marginally different values
  # scale_size_manual(name = "Rat Control", values = c(4, 3)) +
  
  # Remove fill legend and replace the fill legend using the newly created size
  guides(
    col = "none",
    shape = guide_legend(override.aes = list(shape = c(24, 21))),
    fill = guide_legend(override.aes = list(col = c(
      "cornflowerblue", "darkorange"
    )))
  )



# study design plot
# may 2019
#anthony

source("./rcode/davidson_2019_theme.r")
# source("./rcode/figures/study-design-data.R")

p.design <- plot.dat.all1 %>%
  mutate(grid = as.numeric(factor(grid)))

# glimpse(plot.dat.all1)

# adding NA grid to plot

bd.row <- p.design[1, ]
bd.row$grid <- "blank"
bd.row$grid <- "blank"
bd.row$grid <- "blank"
bd.row$grid <- "blank"
bd.row$Date <- NA

# bind row to plot data
p.design145 <- rbind(p.design, bd.row)
# glimpse(p.design145$grid)

# labels
rat.labs <- c("No", "Reduced")

#re-factoring
p.design1 <- p.design145  %>%
  mutate(grid = factor(grid, levels = c(
    "1", "2", "3", "4", "blank", "5", "6", "7", "8"
  )),
  Rats = factor(Rats, labels = rat.labs))


#checking
# tail(p.design1)
# tail(filter(p.design1, grid == "blank"))
# levels(p.design1$grid)
# labels() <- c("1","2","3","4"," ","5","6","7","8")

# p.design1$grid <- factor(p.design1$grid,  labels = c("1","2","3","4"," ","1","2","3","4"))

#plot raw seed
fig.3.plot.seed <- ggplot(p.design1,
                          aes(
                            y = cum.seed,
                            col = Rats,
                            shape = Valley,
                            fill = Control,
                            x = Date
                          ),
                          size = 4) +
  # geom_line(col = "grey50") +
  geom_point(aes(size = Rats,
                 group = grid), stroke = 1.25) +
  
  scale_color_manual(name = "Stoat Control",
                     values = c("white", "black")) +
  
  scale_shape_manual(name = "Ecosystem",
                     values = c(24, 21)) +
  
  # manually define the fill colours
  
  scale_fill_manual(name = "Stoat Control",
                    values = c("cornflowerblue", "darkorange")) +
  # geom_abline(intercept = 5, slope = 0, size = 1) +
  # theme
  theme_new() +
  
  scale_size_manual(name = "Rat Control", values = c(4, 3)) +
  # labels
  # scale_y_discrete(labels = c("1","2","3","4"," ","1","2","3","4")) +
  # scale_x_date() +
  # defining size with 2 marginally different values
  
  # Remove fill legend and replace the fill legend using the newly created size
  guides(
    col = "none",
    size = guide_legend(override.aes = list(shape = c(16, 1))),
    shape = guide_legend(override.aes = list(
      shape = c(24, 21), size = 4
    )),
    fill = guide_legend(override.aes = list(
      col = c("cornflowerblue", "darkorange"),
      size = 4
    ))
  ) +
  #   col = "none",
  #   # size = guide_legend(override.aes = list(shape = c(16, 1))),
  #   shape = guide_legend(override.aes = list(
  #     shape = c(24, 21), size = 4
  #   ))
  #   # fill = guide_legend(override.aes = list(
  #     # col = c("cornflowerblue", "darkorange"), size = 4
  #   # ))
  # ) +
  
  #   geom_hline(yintercept = 0,
#              lty = 5,
#              alpha = 0.7)  +
#
xlab(expression(paste("Time", "(", italic(t), ")"))) +
  
  ylab(expression(paste("Available seed ", "(", italic(Seed[jt]), ")")))

fig.3.plot.seed


# sorting summary dataset for the last time -------------------------------
#summarising seed
table(is.na(p.design1$grid))
#remove NA for making space in last plot
p.design1 <- p.design1[1:144, ]
#no grid na anymore
table(is.na(p.design1$grid))

#making datasest
p.design2 <- p.design1 %>%
  group_by(Control, Valley, Date) %>%
  summarise(cum.seed = mean(cum.seed),
            Rats = factor("Full", levels = c("Full", "Reduced"))) %>%
  ungroup() %>%
  mutate(grid = factor(paste(Control, Valley)))

# grouping of grid correct?
levels(p.design2$grid)
p.design2$Rats

glimpse(p.design2)
#plot summary seed

# Add NA so legend works
row.input.leg <- p.design2[1, ]
row.input.leg$Control <- NA
row.input.leg$Valley <- NA
row.input.leg$Date <- NA
row.input.leg$cum.seed <- NA
row.input.leg$Rats <- "Reduced"
row.input.leg$grid <- NA

p.design3 <- rbind(p.design2, row.input.leg)
glimpse(p.design3)
levels(p.design3$Rats)
col = c("cornflowerblue", "darkorange")

#plot raw seed
fig.3.plot.seed.sum <- ggplot(p.design2,
                              aes(
                                y = cum.seed,
                                col = Rats,
                                shape = Valley,
                                fill = Control,
                                x = Date
                              ),
                              size = 4) +
  geom_line(col = "grey50") +
  geom_point(aes(size = Rats,
                 group = grid), stroke = 1.25) +
  
  scale_color_manual(name = "Stoat Control",
                     values = c("white", "black")) +
  
  scale_shape_manual(name = "Ecosystem",
                     values = c(24, 21)) +
  
  scale_size_manual(name = "Rat Control", values = c(4, 3)) +
  # manually define the fill colours
  
  scale_fill_manual(name = "Stoat Control",
                    values = c("cornflowerblue", "darkorange")) +
  # geom_abline(intercept = 5, slope = 0, size = 1) +
  # theme
  theme_new() +
  
  
  # labels
  # scale_y_discrete(labels = c("1","2","3","4"," ","1","2","3","4")) +
  # scale_x_date() +
  # defining size with 2 marginally different values
  
  # Remove fill legend and replace the fill legend using the newly created size
  guides(
    col = "none",
    size = "none",
    shape = guide_legend(override.aes = list(
      shape = c(24, 21), size = 4
    )),
    fill = guide_legend(override.aes = list(
      col = c("cornflowerblue", "darkorange"),
      size = 4
    ))
  ) +
  #   col = "none",
  #   # size = guide_legend(override.aes = list(shape = c(16, 1))),
  #   shape = guide_legend(override.aes = list(
  #     shape = c(24, 21), size = 4
  #   ))
  #   # fill = guide_legend(override.aes = list(
  #     # col = c("cornflowerblue", "darkorange"), size = 4
  #   # ))
  # ) +
  
  #   geom_hline(yintercept = 0,
#              lty = 5,
#              alpha = 0.7)  +
#
xlab(expression(paste("Time", "(", italic(t), ")"))) +
  
  ylab(expression(paste("Available seed ", "(", italic(Seed[jt]), ")")))

fig.3.plot.seed.sum



# Combine them ------------------------------------------------------------

#plot raw seed
fig.3.seed <- ggplot(p.design1,
                     aes(
                       y = cum.seed,
                       col = Rats,
                       shape = Valley,
                       fill = Control,
                       x = Date
                     )) +
  # geom_line(col = "grey50") +
  geom_point(aes(size = Rats,
                 group = grid),
             stroke = 1.25,
             alpha = 0.3) +
  
  scale_color_manual(name = "Stoat Control",
                     values = c("white", "black", "white")) +
  
  scale_shape_manual(name = "Ecosystem",
                     values = c(24, 21)) +
  
  scale_size_manual(name = "Rat Control", values = c(2.5, 3, 2.5)) +
  # manually define the fill colours
  
  scale_fill_manual(
    name = "Stoat Control",
    values = c("cornflowerblue", "darkorange", "cornflowerblue")
  ) +
  
  
  geom_line(
    data = p.design2,
    aes(y = cum.seed,
        x = Date),
    size = 0.95,
    col = "grey50"
  ) +
  
  geom_point(
    data = p.design2,
    aes(
      y = cum.seed,
      col = Rats,
      shape = Valley,
      fill = Control,
      x = Date
    ),
    size = 7
  ) +
  
  # Remove fill legend and replace the fill legend using the newly created size
  guides(
    col = "none",
    size = guide_legend(override.aes = list(
      shape = c(16, 1), alpha = 1
    )),
    shape = guide_legend(override.aes = list(
      shape = c(24, 21), size = 4
    )),
    fill = guide_legend(override.aes = list(
      col = c("cornflowerblue", "darkorange"),
      size = 4
    ))
  ) +
  
  xlab(expression(paste("Time", "(", italic(t), ")"))) +
  
  ylab(expression(paste("Available seed ", "(", italic(Seed[jt]), ")"))) +
  theme_new()



fig.3.seed

# export plot for example vignette
jpeg("./figs/fig-3.1-seed.jpeg")
fig.3.seed
dev.off()

```





## Resources
