---
title: "Simple models for beech forests"
author: "Anthony Davidson"
date: '2019-05-16'
output:
  html_document:
    df_print: paged
  html_notebook: default
subtitle: For Stoat control in Beech forest paper
bibliography: Beech-forests.bib
type: post
---

```{r global-options, message=FALSE, warning=FALSE, include=FALSE}
# export .r code only
# knitr::purl("./Davidson_2019_BeechForest.Rmd")

# render draft to webpage
# rmarkdown::render(input = "Davidson_2019_BeechForest.Rmd")
# ,
#                   output_format = "html_document",
#                   output_file = "Davidson_2019_t.html")

# compareGroups::cGroupsGUI(mtcars)
#document global rules
knitr::opts_chunk$set(comment=NA,
                      fig.path = "./figs/",
                      echo=FALSE, 
                      fig.height=6, 
                      fig.width=10,
                      message=FALSE, 
                      warning=FALSE)
# how do I do this??
# ,eval = FALSE,include = FALSE


```

# Overview

```{r eniviroment, message=FALSE, warning=FALSE, include=FALSE}
# libraries needed
source("./rcode/r-packages-needed.R", echo = FALSE)
source("./rcode/theme_raw_fig3s.r", echo = FALSE)
source("./rcode/davidson_2019_theme.r", echo = FALSE)

# ("./Davidson_2019_Data_wrangling.Rmd", echo = FALSE)

# not working
# source("./Davidson_2019_Data_wrangling.Rmd", echo = FALSE, )
```

# Descriptive plots

#### Temporal variation in seed

The overall trend in each season is well documented. We have collected data every four months.

```{r seasonal-seed, echo=FALSE, message=FALSE, warning=FALSE}
annual.trend <- ggplot(seed.paper.2, aes(y = seedm2, x = year, fill = valley)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  scale_fill_manual(values=c("#999999", "#E69F00")) +
  labs(x = "Year", y = "Total seed per m2??") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_bw() +
  ggtitle("Annual trend")
  
  
seasonal.trend <- ggplot(seed.paper.2, aes(y = seedm2, x = month, fill = valley)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  scale_fill_manual(values=c("#999999", "#E69F00")) +
  labs(x = "Month", y = "Total seed per m2??") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_bw()+
  ggtitle("Seasonal trend")

gridExtra::grid.arrange(annual.trend, seasonal.trend, ncol = 2)
```

#### Tree `~spp`

But there could be more to this because there are differences between the different types of seed and the two valley systems

```{r beech seed2, echo=FALSE}
# plot data
seed.p1 <- seed %>%
              mutate(grid = paste(valley,grid.id))
                
            
# plot
p1 <- ggplot(seed.p1, aes(y = red, x = trip.no, shape = valley, col = grid, group = grid)) +
           geom_point()+
    geom_line()+ 
  theme(legend.position = "none")

p2 <- ggplot(seed.p1, aes(y = silver, x = trip.no, shape = valley, col = grid, group = grid)) +
           geom_point()+
    geom_line()
    
p3 <- ggplot(seed.p1, aes(y = mountain, x = trip.no, shape = valley, col = grid, group = grid)) +
           geom_point()+
    geom_line() + 
  theme(legend.position = "none")

gridExtra::grid.arrange(p1, p2, p3)
  # seed <- tibble(seed)
# seed.table <- seed[1:3,]
# kable(seed.table)

```

### Spatial variation in seed

We have replicated spatial areas in two ways in our study. We have replicated ecosystems (Valleys) twice and replicated six grids within each one.

### `~valley`

There is a cool bit of code below of  dealing with vaibale plotting without having to use `gather`.

```{r beech seed3, echo=FALSE}
ggplot(seed, aes(y = red, x = trip.no)) +
  geom_point(aes(y = red,color="red"))+
  geom_point(aes(y = silver,color="silver"))+
  geom_point(aes(y = mountain,color="mountain"))+
  stat_summary(color="blue",fun.y="mean",geom="line")+
  facet_grid(~valley) +
  theme_bw()
```

This shows that infact there are differences in the forest composition of the two valleys that have differences in control of stoats.

This is important because differences or lack of differences in the two valleys may simply be due to the different forest compositions rather than the removal of stoats from one system.

### `~grid`

```{r beech seed4, echo=FALSE}
# has to be raw data
ggplot(seed, aes(y = red, x = trip.no)) +
  geom_point(aes(y = red,color="red"))+
  geom_point(aes(y = silver,color="silver"))+
  geom_point(aes(y = mountain,color="mountain"))+
  stat_summary(color="blue",fun.y="mean",geom="line")+
  facet_grid(~grid.id) +
  theme_bw()
```

#### From the finalised dataset

```{r}
ggplot(seed.paper.2, aes(y = seedm2, x = trip)) +
  geom_point()+
  facet_grid(valley~grid.id) +
  theme_bw() +
  ggtitle("Total seedfall")

ggplot(seed.paper.2, aes(y = cum.seed, x = trip)) +
  geom_point()+
  geom_line() +
  facet_grid(valley~grid.id) +
  theme_bw() +
  ggtitle("Total accumulative seedfall")

ggplot(seed.paper.2, aes(y = log.cum.seed, x = trip)) +
  geom_point()+
  geom_line() +
  facet_grid(valley~grid.id) +
  theme_bw() +
  ggtitle("ifelse(cum.seed > 0, log(cum.seed), 0) seedfall")
```

# Discussion

## Bug fixes

**THIS IS the error...** Over the past ... ages I have had issues with factors, characters and merging datasets. Here is where the biggest error in my analysis is found. I have worked through it below (to my shame!)

### Questions still needing asked

1. log of -inf needs to be replaced with zero before analysis?

### Notes

- There is NO red beech recorded in the Hollyford valley

- There is larger peaks of overall food in the Eglinton Valley compared to the Hollyford


# References


