---
title: "Beech forest dynamics"
subtitle: "A simulation"
author: "Anthony Davidson"
date: '2019-05-16'
output:
  word_document: 
    reference_docx: 
      template.docx
bibliography: Beech-forests.bib
---

```{r}
#render this file as follows
# render(rmarkdown::word_document("./Davidson_2019_Simulation.Rmd"))
```

This simple simulation model generates Figure 1 of this publication. This figure explains how I expect population dynamics to 'deterministically work'play-out" given our current understanding of Beech forest dynamics [@choquenot2000; @ruscoe2001; @blackwell2001; @blackwell2003; @ruscoe2005; @tompkins2006 @ tompkins2013; @holland2015;  @latham2017], primarily seed availability (Figure 1).

## Setup

```{r global-options, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(comment=NA,
                      fig.path = "../figs/",
                      echo=FALSE,
                      message=FALSE, 
                      warning=FALSE)

# how do I do this??
# ,eval = FALSE,include = FALSE

library(citr)
# citr::md_cite("Beech-forest.bib")
# bib(file = "Beech-forests.bib")
```

# Overview

```{r eniviroment}
# libraries needed
source("./R/r-packages-needed.R", echo = FALSE)
source("./R/theme_raw_fig3s.r", echo = FALSE)
source("./R/davidson_2019_theme.r", echo = FALSE)
```

New Zealand beech forests exhibit boom-bust dynamics orginally (after many edits looks like this vignette and Figure 1 below.

```{r fig.cap=cap.pred, fig.height=4, fig.width=6}
#data file
source("./R/ecosystem-simulation/sim-raw-data.R", echo = FALSE)

#file code
source("./R/ecosystem-simulation/sim-code-plot.R", echo = FALSE)

#plot
result.plot

cap.pred <- c("Each arrow and label represents a prediction we tested (Prediction B to D). Each prediction represents a collection of previous studies that have suggested how mouse populations may respond to seed availability in the presence or absence of stoats. A) during the years when no seed is available (non-mast years; Panel A); B) at the peak of mouse abundance (winter or spring); C) the season when mice populations are responding rapidly to increasing seed abundance (summer to winter in mast years; Panel B); D) when mouse abundance declines (spring to summer; Panel B).  The top panel represents the average seed availability cycle in New Zealand Native Beech Forests between non-mast (no shading) and mast years (shaded grey). The bottom panel represents the expected response of mouse abundance ($N_{j,t}$) to the variation in seed availability ($Seed_{j,t}) above where solid yellow symbols represent locations where stoats are un-controlled")
```

Where, beech trees mast in spatial synchronised but annually variable years dependant [@wardle1991]. Mice populations have invaded these systems and studies have shown that populations response numerically to changes in resources (beech seed) and mice have been modelled under a range of both functional and numerical responses [@king1983]. 

# Introduction

I have made a very simple simulation of the expected relationships from the literature. *For more detail see thesis drafts [here](https://www.ssnhub.com/phd-thesis/).*

## Season

```{r sim-season}
#simulated data......
#overall variables
season <- rep(as.character(as.factor(c("Summer", "Autumn", "Winter", "Spring"))),3)

# kable(table(season))
```

## Mouse Abundance

```{r sim-abundance, echo=TRUE}
#Abundance
no.stoats <- c(25,15,25,15,10,160,200,200,25,15,25,20)
stoats <- c(12,10,12,5,7,75,100,100,10,3,5,4)

```

## Beech Seed 

```{r sim-seed, echo=TRUE}
#seed
beech.seed <- c(0,0,0,0,0,rnorm(1,1550,1),rnorm(1,1800,1),rnorm(1,2000,1),0,0,0,0)
lcl.seed <- c(0,0,0,0,rnorm(1,10,2),rnorm(1,1000,50),rnorm(1,3000,100),0,0,0,0,0)
ucl.seed <- c(0,0,0,0,rnorm(1,190,10),rnorm(1,3000,50),rnorm(1,5000,100),0,0,0,0,0)
stata <- seq(1,12,1)

```

## Stoat control

```{r echo=TRUE}
#control <- as.factor(c(rep(c("no.stoats"),4),rep(c("stoats"),4)))
control <- factor(rep(1,12))
# kable(table(control))
```

## Date

```{r echo=TRUE}
#date
date <- as.Date(as.character(c("1999-02-01","1999-05-01","1999-08-01","1999-11-01",
                               "2000-02-01","2000-05-01","2000-08-01","2000-11-01",
                               "2001-02-01","2001-05-01","2001-08-01","2001-11-01")))

```

### Labelling

```{r echo=TRUE}
#date
date <- as.Date(as.character(c("1999-02-01","1999-05-01","1999-08-01","1999-11-01",
                               "2000-02-01","2000-05-01","2000-08-01","2000-11-01",
                               "2001-02-01","2001-05-01","2001-08-01","2001-11-01")))
```

## Create dataset

It is always nice to export csv data with meaningful labels.

### Labelling

```{r echo=TRUE}
labels1 <-  c("Summer", "Autumn", "Winter", "Spring", "Summer", "Autumn", "Winter", "Spring")

labels2 <-  c("", "", "Non-mast year", "", "", "", "Mast year", "")

```

## Build dataset

```{r}
#merge data
dat <- data.frame(season,stata,date,stoats,no.stoats,beech.seed,lcl.seed,ucl.seed)
# glimpse(dat)

# reduce data to 10 not 12 seasons
dat <- filter(dat,stata < 11 & stata > 1)
# dat <- filter(dat,stata > 2)


############## hypothesis plot data
#short to long data switch
dat1 <- select(dat,stoats,no.stoats,season, date, stata, beech.seed) %>%
  gather(control,value,stoats:no.stoats)

# str(sim.dat)
# glimpse(dat1)
# kable(head(dat1))
# reduce data to 10 not 12 seasons
# dat1 <- filter(dat1,stata < 11 & stata > 1) %>%
#   droplevels()

```

### Save data

```{r}
# save data
write_csv(dat1, "./data/simulated_data.csv")

#import data for plot
sim.dat <- read_csv(file = "./data/simulated_data.csv")
```

# Results

### Plot data

```{r}
# a cleaner way to deal with code
# source("./rcode_publication/figure_rcode/figure_one_data.R")
# data
sim.dat <- read_csv("./data/simulated_data.csv")
# str(sim.dat)
# glimpse(dat1)
# kable(head(sim.dat))
# reduce data to 10 not 12 seasons
# dat1 <- filter(dat1,stata < 11 & stata > 1) %>%
#   droplevels()

```

### Plot symbols/labels

```{r echo=TRUE}
# build points data
# tibble my life
arrow.length <- 40
touchoff.distance <- 10 # distance between data and start of arrow
arrowhead.size <- 3 # in millimeters
time.loc <- as.character()

# "1999-09-31", "2000-05-31", "2000-07-31", "2000-12-31"

points.dat <- tibble(
  prediction = as.factor(c("A", "C", "B", "D")),
  value = as.numeric(c(15, 88, 108, 60)),
  date = as.Date(c("1999-08-20", "2000-05-01", "2000-09-13", "2000-12-25")))

# c("1999-02-01","1999-05-01","1999-08-01","1999-11-01","2000-02-01","2000-05-01","2000-08-01","2000-11-01", "2001-02-01","2001-05-01","2001-08-01","2001-11-01")

```

### Seed plot

```{r mod-code-seed-new, fig.height=2, fig.width=6}
#plot 
pseed <-ggplot(sim.dat, aes(y = beech.seed, x = date)) +
  geom_rect(aes(xmin=ymd("2000-12-31"),xmax = ymd('1999-12-31'), ymin = -Inf, ymax = Inf), fill = "grey90")+
  geom_line(aes(y = beech.seed, x = date), size = 0.9, colour = "grey80") +
  geom_point(aes(y = beech.seed, x = date, fill = control), fill = "black",stroke = 1, shape = 23, size = 2) +
  xlab("") +
  ylab(expression(paste("Seed ", "(" ,italic(S[jt]),")"))) +
  scale_y_continuous(expand = c(0.05, 0.05)) +
  scale_x_date(breaks = date, date_labels =  labels1, expand = c(0.05, 0.05)) +
  # scale_fill_manual(value ss = c("black")) +
  
  annotate("text", x=as.Date('2000-05-15'), y = 2000,
           label = "Masting year", size = 3.5, colour = "black") +
  annotate("text", x=as.Date('1999-07-30'), y = 2000,
           label = "Non-mast year", size = 3.5, colour = "black") +
  annotate("text", x=as.Date('2001-03-30'), y = 2000,
           label = "Non-mast", size = 3.5, colour = "black") +
  theme_bw()+
  theme_new() +
  theme(plot.margin=margin(t=0, r = 0, l = 0, 0, unit="cm"),

        # plot.margin = unit(c(0.1,0.1,0.1,0.1), units = "cm"),

        legend.position = "none",
        legend.key = element_blank(),
        legend.background = element_rect(fill="white", size=1),
        legend.key.size=unit(1,"cm"),
        legend.text = element_text(colour = "black", size =12),
        legend.title = element_text(colour = "black", size =12),

        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(2, "lines"),
        panel.border = element_blank(),
        axis.title.y = element_text(colour = "black",size =12,angle = 90),
        axis.title.x = element_text(colour = "black",size = 10),
        axis.text.y=element_text(colour = "black",size = 10),
        axis.text.x = element_blank(),

        axis.ticks.x = element_line(size = 1),
        axis.ticks.y = element_line(size = 1),
        axis.line.x = element_line(size = 1),
        axis.line.y = element_line(size = 1))

pseed

# orginal plot
pseed1 <-   pseed +  
  # geom_abline(slope = 0, intercept = 0, lty = 3, alpha = 0.8) +
  theme(plot.margin=margin(t=0.03, r = 0, l = 0, b = -0.03, unit="cm"),

        # plot.margin = unit(c(0.1,0.1,0.1,0.1), units = "cm"),

        legend.position = "none",
        legend.key = element_blank(),
        legend.background = element_rect(fill="white", size=1),
        legend.key.size=unit(1,"cm"),
        legend.text = element_text(colour = "black", size =12),
        legend.title = element_text(colour = "black", size =12),

        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(2, "lines"),
        panel.border = element_blank(),
        axis.title.y = element_text(colour = "black",size =12,angle = 90),
        axis.title.x = element_blank(),
        axis.text.y=element_text(colour = "black",size = 10),
        axis.text.x = element_blank(),

        axis.ticks.x = element_blank(),
        axis.ticks.y = element_line(size = 1),
        axis.line.x = element_blank(),
        axis.line.y = element_line(size = 1))
  
pseed1
#save
# png("../figs/fig-1-seed.png")
# pseed
# dev.off()
```

### Mouse plot

```{r fig.height=4, fig.width=6}
pmice <- ggplot(sim.dat, aes(y = value, x = date)) +
  geom_rect(aes(xmin=ymd("2000-12-31"),xmax = ymd('1999-12-31'), ymin = -Inf, ymax = Inf), fill = "grey90") +
  
  geom_line(aes(group = control, col = control), size = 0.9) +
  geom_point(aes(fill = control, colour = control), stroke = 1, shape = 23, size = 2, alpha = 0.8) +
  
  xlab("Time (t)") + 
  ylab(expression(paste("Mice"," ", "(" ,italic(N[jt]),")"))) +
  
  scale_y_continuous(expand = c(0.05, 0.05)) +
  scale_x_date(breaks = date, date_labels =  labels1, expand = c(0.05, 0.05)) +

    geom_segment(data = points.dat, aes(x = date, y = value,
                               xend = date, yend = value + touchoff.distance + arrow.length),
                 arrow = arrow(length = unit(arrowhead.size, "mm"), ends = "first"), size = 1.25, colour = "red") +

  geom_label(data = points.dat, aes(x = date, y = value, label = prediction),
             nudge_x = 20,
             nudge_y = 40)  +
    scale_colour_manual(name = "Stoat control",
                      labels = c("Yes", "No"),
                      values = c("cornflowerblue","darkorange")) +

  scale_fill_manual(name = "Stoat control",
                    labels = c("Yes", "No"),
                    values = c("cornflowerblue","darkorange")) +
  theme_bw()+
  theme_new() +
  theme(
    # plot.margin=margin(t=0, r = 0, l = 0, 0, unit="cm"),

        # plot.margin = unit(c(0.1,0.1,0.1,0.1), units = "cm"),

        legend.position = "none",
        legend.key = element_blank(),
        legend.background = element_rect(fill="white", size=1),
        legend.key.size=unit(1,"cm"),
        legend.text = element_text(colour = "black", size =12),
        legend.title = element_text(colour = "black", size =12),

        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(2, "lines"),
        panel.border = element_blank(),
        axis.title.y = element_text(colour = "black",size =12,angle = 90),
        axis.title.x = element_text(colour = "black",size = 10),
        axis.text.y=element_text(colour = "black",size = 10),
        axis.text.x = element_text(colour = "black",size = 10),

        axis.ticks.x = element_line(size = 1),
        axis.ticks.y = element_line(size = 1),
        axis.line.x = element_line(size = 1),
        axis.line.y = element_line(size = 1))
pmice

#save
png("../figs/fig-1-seed.png")
pmice
dev.off()
```

### Joint plot

```{r fig.height=4, fig.width=6}
# makes graphs smaller
result.plot2 <- cowplot::plot_grid(pseed1, pmice ,nrow = 2, 
                                    align="v",
                                    rel_heights = c(0.7, 1.3))

result.plot2
# plot code
# , 
                                    # labels = c("a)", "b)")

# Save plot

png("../figs/fig-1-all.png")
result.plot2
dev.off()
```

#### Old plot

```{r orginal-plot}
#plot 
pseed <- ggplot(sim.dat, aes(y = beech.seed, x = date)) +
  geom_rect(aes(xmin=ymd("2000-12-31"),xmax = ymd('1999-12-31'), ymin = -Inf, ymax = Inf), fill = "grey90")+
  geom_line(aes(y = beech.seed, x = date), size = 1.1) +
  geom_point(aes(y = beech.seed, x = date, fill = control), fill = "black",stroke = 1.5, shape = "square", size = 6) +
  xlab("") +
  ylab(expression(paste("Seed ", "(" ,italic(S[jt]),")"))) +
  
  scale_y_continuous(expand = c(0.05, 0.05)) +
  scale_x_date(breaks = date, date_labels =  labels1, expand = c(0.05, 0.05)) +
  # scale_fill_manual(value ss = c("black")) +
  
  annotate("text", x=as.Date('2000-05-15'), y = 2000,
           label = "Masting year", size = 4, colour = "black", family = "Times") +
  annotate("text", x=as.Date('1999-07-30'), y = 2000,
           label = "Non-mast year", size = 4, colour = "black", family = "Times") +
  annotate("text", x=as.Date('2001-03-30'), y = 2000,
           label = "Non-mast", size = 4, colour = "black", family = "Times") +
  
  theme_new() +
  theme(
    
    # strip.background = element_blank(),
    #     strip.text.y = element_blank(),
    #     plot.title = element_text(hjust = 0, size=14, family = "Times", color="black"),

        plot.margin = unit(c(0.1,0.1,0.1,0.1), units = "cm"),

        # plot.margin=margin(t=0.5, r = 0, l = 0, -0.05, unit="cm"),
        # legend.position = "none",
        # legend.key = element_blank(),
        # legend.background = element_rect(fill="white", size=1),
        # legend.key.size=unit(1,"cm"),
        # legend.text = element_text(colour = "black", size =12, family = "Times"),
        # legend.title = element_text(colour = "black", size =12, family = "Times"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        # panel.spacing = unit(2, "lines"),
        panel.border = element_blank(),
        # axis.title.y = element_text(colour = "black",size =12, family = "Times", angle = 90),
        axis.text.y=element_text(colour = "black",size = 14, family = "Times"),

        axis.title.x = element_blank(),
        axis.text.x = element_blank(),

        axis.ticks.x = element_line(size = 1),
        axis.line.x = element_line(size = 1),

        axis.ticks.y = element_line(size = 1),

        axis.line.y = element_line(size = 1))


# pseed


# build points data
# tibble my life
arrow.length <- 10
touchoff.distance <- 10 # distance between data and start of arrow
arrowhead.size <- 3 # in millimeters
time.loc <- as.character()

# "1999-09-31", "2000-05-31", "2000-07-31", "2000-12-31"

points.dat <- tibble(
  prediction = as.factor(c("A", "C", "B", "D")),
  value = as.numeric(c(15, 88, 108, 60)),
  date = as.Date(c("1999-08-20", "2000-05-01", "2000-09-13", "2000-12-25")))

# c("1999-02-01","1999-05-01","1999-08-01","1999-11-01","2000-02-01","2000-05-01","2000-08-01","2000-11-01", "2001-02-01","2001-05-01","2001-08-01","2001-11-01")

pmice <- ggplot(sim.dat, aes(y = value, x = date)) +
  geom_rect(aes(xmin=ymd("2000-12-31"),xmax = ymd('1999-12-31'), ymin = -Inf, ymax = Inf), fill = "grey90") +
  
  geom_line(aes(group = control, col = control), size = 1.1) +
  geom_point(aes(fill = control, colour = control), stroke = 1.5, shape = "square", size = 4.5) +
  
  xlab("Time (t)") + 
  ylab(expression(paste("Mice"," ", "(" ,italic(N[jt]),")"))) +
  
  scale_y_continuous(expand = c(0.05, 0.05)) +
  scale_x_date(breaks = date, date_labels =  labels1, expand = c(0.05, 0.05)) +
  
  # scale_fill_manual(values = c("white", "black")) +
  
  scale_colour_manual(name = "Stoat control",
                      labels = c("Yes", "No"),
                      values = c("cornflowerblue","darkorange")) +

  scale_fill_manual(name = "Stoat control",
                    labels = c("Yes", "No"),
                    values = c("cornflowerblue","darkorange")) +
  
  
  geom_segment(data = points.dat, aes(x = date, y = value,
                               xend = date, yend = value + touchoff.distance + arrow.length),
                 arrow = arrow(length = unit(arrowhead.size, "mm"), ends = "first"), size = 1.25, colour = "red") +
  
  geom_label(data = points.dat, aes(x = date, y = value, label = prediction),
             nudge_x = 10,
             nudge_y = 30) +
  
  theme_new()+ 
  theme(legend.position = c(0.1,0.5),

        # plot.margin=margin(t=0, r = 0, l = 0, 0, unit="cm"),

        # plot.margin = unit(c(0.1,0.1,0.1,0.1), units = "cm"),

        # # legend.position = "none",
        # legend.key = element_blank(),
        # legend.background = element_rect(fill="white", size=1),
        # legend.key.size=unit(1,"cm"),
        # legend.text = element_text(colour = "black", size =16, family = "Times"),
        # legend.title = element_text(colour = "black", size =16, family = "Times"),
        
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(2, "lines"),
        panel.border = element_blank(),

        axis.title.y = element_text(colour = "black",size =14, family = "Times", angle = 90, vjust =  1),
        axis.title.x = element_blank(),
        axis.text.y=element_text(colour = "black",size = 14, family = "Times"),
        axis.text.x = element_text(colour = "black",size = 12, family = "Times", vjust = 1),

        axis.ticks.x = element_line(size = 1),
        axis.ticks.y = element_line(size = 1),
        axis.line.x = element_line(size = 1),
        axis.line.y = element_line(size = 1),

        strip.text = element_text(face="bold",colour = "black",size =14, family = "Times"))

# pmice

# makes graphs smaller
  result.plot <- cowplot::plot_grid(pseed, pmice ,nrow = 2, 
                                    align="v", 
                                    labels = c("a)", "b)"),
                                    rel_heights = c(0.7, 1.3))

 result.plot

# Save plot

png("../figs/fig-1-1.png")
result.plot
dev.off()
```



# Discussion

### Saving

```{r message=FALSE, warning=FALSE}
# export data
### ????

# export plot for example vignette
png("../figs/fig-1-3.png")
result.plot
dev.off()
```

# Appendix 1
## Old code 

```{r data-generation-old, include=FALSE, echo = TRUE}
#simulated data......
#overall variables
season <- rep(as.character(as.factor(c("Summer", "Autumn", "Winter", "Spring"))),3)

#Abundance
no.stoats <- c(25,15,25,15,10,160,200,200,25,15,25,20)
stoats <- c(12,10,12,5,7,75,100,100,10,3,5,4)

#seed
beech.seed <- c(0,0,0,0,0,rnorm(1,1550,1),rnorm(1,1800,1),rnorm(1,2000,1),0,0,0,0)
lcl.seed <- c(0,0,0,0,rnorm(1,10,2),rnorm(1,1000,50),rnorm(1,3000,100),0,0,0,0,0)
ucl.seed <- c(0,0,0,0,rnorm(1,190,10),rnorm(1,3000,50),rnorm(1,5000,100),0,0,0,0,0)
stata <- seq(1,12,1)
control <- factor(rep(1,12))
#control <- as.factor(c(rep(c("no.stoats"),4),rep(c("stoats"),4)))
labels1 <-  c("Summer", "Autumn", "Winter", "Spring", "Summer", "Autumn", "Winter", "Spring")

#date
date <- as.Date(as.character(c("1999-02-01","1999-05-01","1999-08-01","1999-11-01",
                               "2000-02-01","2000-05-01","2000-08-01","2000-11-01",
                               "2001-02-01","2001-05-01","2001-08-01","2001-11-01")))

#merge data
dat <- data.frame(control,season,stata,date,stoats,no.stoats,beech.seed,lcl.seed,ucl.seed)
# glimpse(dat)

# change levels
levels(dat$season) = c("Summer", "Autumn", "Winter", "Spring")

# reduce data to 10 not 12 seasons
dat <- filter(dat,stata < 11 & stata > 1)
# dat <- filter(dat,stata > 2)


############## hypothesis plot data
#short to long data switch

dat1 <- select(dat,stoats,no.stoats,season, date, stata, beech.seed) %>%
  gather(control,value,stoats:no.stoats)

# glimpse(dat1)

# reduce data to 10 not 12 seasons
dat1 <- filter(dat1,stata < 11 & stata > 1) %>%
  droplevels()

# dat <- filter(dat,stata > 2)

write_csv(dat1, "./Data/simulated_data.csv")
sim.dat <- read_csv(file = "./Data/simulated_data.csv")
```

# References


